### 2020-05-26 serial.tex
comment on eq (3)

\eref{backward} suggests that backward delay distributions change over time even if their corresponding forward delay distribution does not change:
Backward delay distributions depend on changes in the primary cohort size over time due to conditionality of observations (i.e., they are measured from a cohort of individuals whose secondary events occurred at the same time) \bmb{I think this is very important and deserves a few more words to clarify, e.g. ``i.e., conditioning on individuals whose secondary events have already occurred means that we tend not to observe longer inter-event times, so that the distribution is truncated [censored??]'' (or something like that)?}.
When the cohort size is increasing (or decreasing) over time, we are biased to observe shorter (or longer) delays when we look backward.
In particular, when the incidence is growing exponentially, we can calculate the amount of bias exactly.
Assuming that the forward delay distribution remains constant during the exponential growth phase ($f_p(\tau) \approx f_0(\tau)$), we can substitute $\psize(t) = \psize(0) \exp(rt)$ in \eref{backward} to obtain:
\begin{equation}
b_0(\tau) \propto \exp(-r\tau) f_0(\tau),
\label{eq:backexp}
\end{equation}
where $r$ is the exponential growth rate.
For faster epidemics (higher $r$), the backward delay distribution will be shorter.
\bmb{I'm not sure whether ``distribution'' should be singular or plural in general.  Singular reads better to me most of the time, but do whatever you want as long as it's consistent}
Changes in the backward delay distribution are also by definition affected by the changes in the forward delay distribution (\eref{backward}).
Therefore, the backward delay distribution will generally differ from the forward delay distribution even if we are measuring time delays that are intrinsic to the life history of a disease (e.g., the incubation period).
These ideas apply to all epidemiological delay distributions and generalize the work by \cite{champredon2015intrinsic} who compared forward and backward generation-interval distributions to describe the realized generation intervals from the perspective of an infector and an infectee, respectively, as well as the work by \cite{britton2019estimation} who showed that \eref{backexp} holds for the backward generation-interval distributions.

>>>>

\eref{backward} suggests that backward delay distributions change over time even if their corresponding forward delay distribution does not change:
Backward delay distributions depend on changes in the primary cohort size over time due to conditionality of observations (i.e., they are measured from a cohort of individuals whose secondary events occurred at the same time) \bmb{I think this is very important and deserves a few more words to clarify, e.g. ``i.e., conditioning on individuals whose secondary events have already occurred means that we tend not to observe longer inter-event times, so that the distribution is truncated [censored??]'' (or something like that)?}.
When the cohort size is increasing (or decreasing) over time, we are biased to observe shorter (or longer) delays when we look backward.
In particular, when the incidence is growing exponentially, we can calculate the amount of bias exactly.
Assuming that the forward delay distribution remains constant during the exponential growth phase ($f_p(\tau) \approx f_0(\tau)$), we can substitute $\psize(t) = \psize(0) \exp(rt)$ in \eref{backward} to obtain:
\begin{equation}
b_0(\tau) \propto \exp(-r\tau) f_0(\tau),
\label{eq:backexp}
\end{equation}
\djde{I do not find it helpful to write $\propto$.  It would be clearer to write it out exactly with the part that doesn't depend on $\tau$ in $[]$.  Then you can follow with ``where $r$ is the exponential growth rate and the expression in square brackets is a constant since it does not depend on $\tau$.''  I think it is helpful to see what the constant is and not make the reader think it through.  Alternatively, you could write the equality and then have your $\propto$ and then say afterwards that we write $\propto$ for convenience since the expression in $[]$ is constant wrt $\tau$.}
where $r$ is the exponential growth rate.
For faster epidemics (higher $r$), the backward delay distribution will be shorter.
\bmb{I'm not sure whether ``distribution'' should be singular or plural in general.  Singular reads better to me most of the time, but do whatever you want as long as it's consistent}
Changes in the backward delay distribution are also by definition affected by the changes in the forward delay distribution (\eref{backward}).
Therefore, the backward delay distribution will generally differ from the forward delay distribution even if we are measuring time delays that are intrinsic to the life history of a disease (e.g., the incubation period).
These ideas apply to all epidemiological delay distributions and generalize the work by \cite{champredon2015intrinsic} who compared forward and backward generation-interval distributions to describe the realized generation intervals from the perspective of an infector and an infectee, respectively, as well as the work by \cite{britton2019estimation} who showed that \eref{backexp} holds for the backward generation-interval distributions.
